---
title: "CreateMailingList"
author: "Bruce Cottman"
date: "January 14, 2015"
output: html_document
---

This report creates the Latino mailing list and various diagnostic graphs.

```{r, echo=FALSE}
### Analyze Last Sale For a given Listing
### 12/15/2014
### Bruce Cottman
###
###
rm(list = ls(all = TRUE)) #CLEAR WORKSPACE
#
invisible(require(XLConnect))
#install.packages("xlsx")
invisible(require(xlsx))
invisible(require(stringr))
#install.packages("dplyr")
#invisible(require(dplyr))
invisible(require(ggplot2))
#install.packages("lubridate")
invisible(require(lubridate))
#install.packages("bitops")
#invisible(require(bitops))
#invisible(require(RCurl))
#install.packages("scales")
#invisible(require(scales))
#install.packages("ggthemes")
#invisible(library(ggthemes))

#
### graphs theme functions
#theme
theme_slanted_axis = function(){
  theme(plot.title=element_text(size=12,face="bold", vjust=2)) +
    theme(axis.text.x = element_text(angle = 50, vjust=1.0, hjust = 1,face="bold",colour = "black")) +
    theme(axis.text.y = element_text(face="bold",colour = "black")) + 
    theme(axis.text=element_text(size=10),axis.title=element_text(size=10,face="bold"))
}
# DB are usually binary and internal use
readDB <- function(wd,fname){
  setwd(wd)
  cat("Loading database file: ",fname,"\n")
  readRDS(fname)
}

ACPARecords = readDB("~/Documents/Real.Estate/input/DB/","DB.APCA.11012015.rds")
ACPARecords = ACPARecords[ACPARecords$paddr != "" ,]
ACPARecords = unique(ACPARecords)
#table(ACPARecords$PUse)
# focus on homes, filter out land, vacancies, etc.
pull = c("SINGLE FAMILY",
        "MULTIFAMILY",
        "MISC. RESIDENCE",
        "MOBILE HOME",
        "CONDOMINIUM"
)
ACPARecords = unique(subset(ACPARecords,ACPARecords$PUse %in% pull ))
cat("Loaded ACPA Records:",nrow(ACPARecords),"\n")

# 

LatinoNames = readDB("~/Documents/Real.Estate/input/DB/","DB.LatinoNames.21092014.rds")
cat("Loaded LATINO last names DB:",nrow(LatinoNames),"\n")
# Latinos in ACPARecords
LatinoInACPARecords = ACPARecords[(ACPARecords$LastName %in% LatinoNames$LastName),]

# Listings of CITIES OF alachua county
readListingDB <- function(wdf,fname,wdl){
  setwd(wdf)
  CityList = read.xlsx(fname,sheetIndex=1,stringsAsFactors = F)
  CityList = CityList[complete.cases(CityList),]
  # Read in  listings
  Start =1
  End = nrow(CityList)
  seqi = as.numeric(seq(Start,End))
# ii=seqi[Stat]
  INITFLAG = 0
  for(ii in seqi){
    if(CityList$FILENAME[ii] != "NONE"){
      setwd(wdl)
      prop=readRDS(CityList$FILENAME[ii])
      # INITIALZE
      if(INITFLAG == 0){
        TCityListings = subset(prop,prop$street %in% "0xxxxxyyyzzz")
        INITFLAG = 1
      }
      cat("Number of Listing Records:",nrow(prop)," for City: ",CityList$CITY[ii],"\n")
      # add to total
      TCityListings = rbind(TCityListings,prop)
    } # END IF
  } #END FOR LOOP
  cat("Total Number of Listing Records:",nrow(TCityListings),"\n")
  TCityListings = unique(TCityListings)
  cat("Total Number of UNIQUE Listing Records:",nrow(TCityListings),"\n")
  TCityListings
}

TCityListings = readListingDB("~/Dropbox/Public/Real.Estate/code","AlachuaCountyCityListingsList.xlsx","~/Documents/Real.Estate/input/DB/")
cat("Loaded Listings DB:",nrow(TCityListings),"\n")
# no land
TCityListings = unique(subset(TCityListings,TCityListings$type != "LAND"))

# listing address cleanup
TCityListings = TCityListings[TCityListings$street != "100 SLIME ST NO STREET ADDRESS" ,]
cat("Eliminate -100 SLIME ST NO STREET ADDRESS - Records Remaining: ",nrow(TCityListings),"\n")
TCityListings$street = str_replace(TCityListings$street,"TERRACE","TER")
TCityListings$street = str_replace(TCityListings$street," LN"," LANE")
TCityListings$street = str_replace(TCityListings$street," ROAD"," RD")
TCityListings$street = str_replace(TCityListings$street," PLACE"," PL")
TCityListings$street = str_replace(TCityListings$street," BOULEVARD"," BLVD")
TCityListings$street = str_replace(TCityListings$street," # A-"," A")
TCityListings$street = str_replace(TCityListings$street," # B-"," B")

TCityListings$street = str_replace(TCityListings$street," # "," APT ")
TCityListings$street = str_replace(TCityListings$street," AVE AVE"," AVE")
TCityListings$street = str_replace(TCityListings$street," UNIT "," - UNIT ")
TCityListings$street = str_replace(TCityListings$street," SW 45","  SW STATE RD 45")
TCityListings$street = str_replace(TCityListings$street," SW 346","  SW COUNTY RD 346")
TCityListings = subset(TCityListings,str_detect(TCityListings$street,"[0-9]{1,} [NW,NE,SE,SW]"))

cat("Eliminate - NO STREET ADDRESS number - Records Remaining: ",nrow(TCityListings),"\n")




# eliminate record columns don't/need in mailing or downstram plots
LatinoInACPARecords = within(LatinoInACPARecords,{FirstName1 = paste0(FirstName1," & ",FirstName2)})
LatinoInACPARecords$FirstName2 = NULL
LatinoInACPARecords = within(LatinoInACPARecords,{Addr1 = paste0(Addr1,", ",Addr2)})
LatinoInACPARecords$Addr2 = NULL
LatinoInACPARecords$Section = NULL
LatinoInACPARecords$Township = NULL
LatinoInACPARecords$Range = NULL
LatinoInACPARecords$MillageCode = NULL
LatinoInACPARecords$SqFt = NULL
#LatinoInACPARecords$HtdSqFt = NULL
LatinoInACPARecords$Acres = NULL
LatinoInACPARecords$Subdivision = NULL
LatinoInACPARecords$AgLandVal = NULL
LatinoInACPARecords$MrktLandVal = NULL
LatinoInACPARecords$BldgVal = NULL
LatinoInACPARecords$MiscVal = NULL
LatinoInACPARecords$SOHDeferAmt = NULL
LatinoInACPARecords$ExemptAmt = NULL
#LatinoInACPARecords$TaxAmt = NULL
LatinoInACPARecords$QualUnqual = NULL
LatinoInACPARecords$SaleInstrument = NULL
#LatinoInACPARecords$LocAddr = NULL
#LatinoInACPARecords$VacImp = NULL
LatinoInACPARecords$Value = NULL
LatinoInACPARecords$freqUseNbhd = 0
# frequency by use by neighborhood
x = as.data.frame(table(LatinoInACPARecords$Neighborhood,LatinoInACPARecords$PUse))
names(x) = c("Neighborhood","Use","Frequency")
x = x[order(x$Use,x$Frequency),]
print(x)
# there must be easier way but this works
# returns list of freqUseNbhd of length of target dataframe
# this is total of both owners and investors
freaky = function (z) {
  i = which(LatinoInACPARecords$Parcel==z) # row index of dataframe
  LatinoInACPARecords$freqUseNbhd[i] = x$Frequency[x$Use == LatinoInACPARecords$PUse[i]
              & x$Neighborhood == LatinoInACPARecords$Neighborhood[i]]
  }

LatinoInACPARecords$freqUseNbhd = sapply(LatinoInACPARecords$Parcel,freaky)
LatinoInACPARecords = with(LatinoInACPARecords,LatinoInACPARecords[order(PUse,Neighborhood,freqUseNbhd),])
LatinoInACPARecords[LatinoInACPARecords$LastName == "DIAZ",]
print("broadmoor test")
cat("Total: ",nrow(LatinoInACPARecords[LatinoInACPARecords$Neighborhood == "BROADMOOR",]),"\n")
cat("Owners: ",nrow(with(LatinoInACPARecords,LatinoInACPARecords[Neighborhood == "BROADMOOR" & paddr == maddr,])),"\n")
cat("Investers: ",nrow(with(LatinoInACPARecords,LatinoInACPARecords[Neighborhood == "BROADMOOR" & paddr != maddr,])),"\n")

LatinoInACPARecords[LatinoInACPARecords$Neighborhood == "BROADMOOR",]
## Eliminate Latinos with li"Neighborhood"stings
# first find Latino owners with listings
LatinoOwners = LatinoInACPARecords[LatinoInACPARecords$paddr == LatinoInACPARecords$maddr,]

LatinoOwnersInListing = intersect(LatinoOwners$paddr,TCityListings$street)
LatinoOwnersNotInListing = LatinoOwners[!(LatinoOwners$paddr %in% LatinoOwnersInListing),]
LatinoOwnersNotInListing$paddr = NULL
LatinoOwnersNotInListing$maddr = NULL
cat("Total Number of Owner Adresses Not in Listing Records:",nrow(LatinoOwnersNotInListing),"\n")
# write out owner mailing list
setwd("~/Dropbox/realestate.output/sales/")
fln = paste0("Latino-OwnInv.MAILING.07022015",".xls")

writeWorksheetToFile(file = fln, data = LatinoOwnersNotInListing, sheet = "Owner")

# next find Latino invetors with listings - none!
LatinoInvestors =subset(LatinoInACPARecords,(LatinoInACPARecords$paddr != LatinoInACPARecords$maddr ))
LatinoInvestorsNotInListing = LatinoInvestors
LatinoInvestorsNotInListing$paddr = NULL
LatinoInvestorsNotInListing$maddr = NULL
cat("Total Number of Investor Adresses Not in Listing Records:",nrow(LatinoInvestorsNotInListing),"\n")

writeWorksheetToFile(file = fln, data = LatinoInvestorsNotInListing, sheet = "Investors")
#
# next find Latino invetors with listings - none!
# this is wrong if other cities than Gainesville
LatinoRentersNotInListing = LatinoInvestorsNotInListing[
  LatinoInvestorsNotInListing$VacImp == "IMPROVED", ]
LatinoRentersNotInListing$LastName = "Current Resident"
LatinoRentersNotInListing$FirstName1 = ""
LatinoRentersNotInListing$Addr1 = LatinoRentersNotInListing$LocAddr
LatinoRentersNotInListing$City = "GAINESVILLE"
LatinoRentersNotInListing$State = "FL"
LatinoRentersNotInListing$Zip = ""
cat("Total Number of Renter Adresses Not in Listing Records:",nrow(LatinoRentersNotInListing),"\n")

writeWorksheetToFile(file = fln, data = LatinoRentersNotInListing, sheet = "Renters")
#
print(getwd())
cat("Saving Latino mailing list to file: ",fln,"\n")
### Graphs
cat("Listing by Usage Type:","\n")
tann = table(TCityListings$type)
#tann
round(prop.table(tann)*100,1)
#
cat("Latino Owners not Listing Usage Type:","\n")
tann = table(LatinoOwnersNotInListing$PUse)
#tann
round(prop.table(tann)*100,1)
cat("Latino Investors not Listing Usage Type:","\n")
tann = table(LatinoInvestorsNotInListing$PUse)
#tann
round(prop.table(tann)*100,1)


#
cat("Latino Owners live in These Cities:","\n")
tann = sort(table(LatinoOwnersNotInListing$City),decreasing=TRUE)
bw = 10
leadingCities = head(tann,n=bw)
leadingCities
#
cat("Latino Investors live in These Cities:","\n")
tann = sort(table(LatinoInvestorsNotInListing$City),decreasing=TRUE)
bw = 10
leadingCities = head(tann,n=bw)
leadingCities

#
bw = 1
yl = "Number in Zip Code"

#png(filename="Listingbyzip.png", width=11*resol, height=8*resol,res=resol)
p = ggplot(TCityListings,aes(x=zip)) 
#p +  geom_histogram(binwidth = bw,aes(fill = ..count..)) +  
  p +  geom_bar(binwidth = bw,aes(fill = ..count..)) +  
  labs(title = "Listing zips",x="Zip Code",y=yl) +
#  scale_fill_gradient(yl, low = "red", high = "green") +
  stat_bin(binwidth=bw, geom="text",size=3, aes(label=..count..), vjust=-.5) +
#  scale_y_continuous(limits = c(20, maxi)) +
  theme_slanted_axis() 
#dev.off()
resol = 600
#listings by broker
setwd("~/Documents/Real.Estate/output/graphs/")
yl = "Number of Listings"
bw = 10
tab = sort(table(TCityListings$broker),decreasing=TRUE)
leadingBrokers = head(tab,n=bw)
leadingBrokers
TopBrokerNames = names(leadingBrokers)
maxi = max(tab)
BListings = subset(TCityListings,TCityListings$broker %in% TopBrokerNames )

## set the levels in order we want
BListings = within(BListings, {broker = factor(BListings$broker, 
                              levels=names(sort(table(BListings$broker),decreasing=TRUE)))})

cat("Broker Listing since 2004 \n")
tabb = table(BListings$broker)
round(prop.table(tabb),2)
#png(filename="BrokerListings.png", width=11*resol, height=8*resol,res=resol)
p = ggplot(BListings,aes(x=broker)) 
#p +  geom_histogram(binwidth = bw,aes(fill = ..count..)) +  
  p +  geom_bar(binwidth = bw,aes(fill = ..count..)) +  
  labs(title = "Broker Listings of Archer, Gainesville, Newberry",x="Broker",y=yl) +
#  scale_fill_gradient(yl, low = "red", high = "green") +
  stat_bin(binwidth=bw, geom="text",size=3, aes(label=..count..), vjust=-.5) +
#  scale_y_continuous(limits = c(20, maxi)) +
  theme_slanted_axis() 
#dev.off()
#quarterly Sales since 2010
SalesOI = subset(ACPARecords,(ACPARecords$SaleDate > as.Date("2000-01-01") & ACPARecords$SaleDate < as.Date("2015-01-01") ))

setwd("~/Documents/Real.Estate/output/graphs/")
yl = "Number of Sales"
bw = 365/4
cat("Quarterly Sales since 2004 \n")
#png(filename="SalesByQuarter.png", width=8*resol, height=8*resol,res=resol)
p = ggplot(SalesOI,aes(x=as.Date(SaleDate))) 
p +  geom_histogram(binwidth = bw,aes(fill = ..count..)) +  
      labs(title = "Quarterly Sales Alachua County",x="Quarter of Sale",y=yl) +
        scale_fill_gradient("Count", low = "red", high = "green") +
        stat_bin(binwidth=bw, geom="text",size=3, aes(label=..count..), vjust=-.5) +
        theme_slanted_axis() 
#dev.off()
# yearly for years gt 2000
cat("Yearly Sales\n")
yl = "Number of Sales"
bw = 365

#png(filename="SalesByYear.png", width=8*resol, height=8*resol,res=resol)
p = ggplot(SalesOI,aes(x=as.Date(SaleDate))) 
p +  geom_histogram(binwidth = bw,aes(fill = ..count..)) +  
        labs(title = "Yearly Sales Alachua County",x="Year of Sale",y=yl) +
        scale_fill_gradient("Count", low = "red", high = "green") +
        stat_bin(binwidth=bw, geom="text",size=3, aes(label=..count..), vjust=-.5) +
        theme_slanted_axis()
#dev.off()
#
cat("Latino Owners not in listings Sales Date Catergories by months\n")
tabb = table(LatinoOwnersNotInListing$SaleDateCategoryLabels)
print(prop.table(tabb))
cat("Percentage of Latino owners not in listings  Records with Sales Date > 2002,",round(sum(LatinoOwnersNotInListing$SaleDate > as.Date("2003-01-01"))
      / nrow(LatinoOwnersNotInListing),2)*100,"\\%\n")
#
#

SalesOI = subset(LatinoOwnersNotInListing,LatinoOwnersNotInListing$SaleDate > as.Date("2000-01-01"))
Adresses = SalesOI
cat("Past Sales of not in Listings\n")
yl = "Number of Past Sales"
bw = 365

#png(filename="ListingPastSalesByYear.png", width=8*resol, height=8*resol,res=resol)
p = ggplot(Adresses,aes(x=as.Date(SaleDate))) 
p +  geom_histogram(binwidth = bw,aes(fill = ..count..)) +  
  labs(title = "Yearly Latino Owner not in Listings Past Sales Alachua County",x="Year of Sale",y=yl) +
  scale_fill_gradient(yl, low = "red", high = "green") +
  stat_bin(binwidth=bw, geom="text",size=3, aes(label=..count..), vjust=-.5) +  
  scale_size_continuous(range = c(3, 8)) +
  theme_slanted_axis()
#dev.off()
cat("Past Sales of  Latino Owner not in Listings by Quarter\n")
yl = "Number of Past Sales"
bw = 365/4

#png(filename="ListingPastSalesByQuarter.png", width=8*resol, height=8*resol,res=resol)
p = ggplot(Adresses,aes(x=as.Date(SaleDate))) 
p +  geom_histogram(binwidth = bw,aes(fill = ..count..)) +  
  labs(title = "Quarterly  Latino Owner not in Listings Past Sales Alachua County",x="Year of Sale",y=yl) +
  scale_fill_gradient(yl, low = "red", high = "green") +
  stat_bin(binwidth=bw, geom="text",size=3, aes(label=..count..), vjust=-.5) +  
  scale_size_continuous(range = c(3, 8)) +
  theme_slanted_axis()
#dev.off()
cat("Owners Sales Price Raw fit ","\n")
    ggplot(LatinoOwnersNotInListing, aes(x = TaxAmt, y = SalePrice)) +
            geom_smooth(method = "lm") + geom_point() +
            scale_x_continuous(expand = c(0, 0)) + 
            scale_y_continuous(expand = c(0, 0))
    lf0 = lm(SalePrice ~ 0 + TaxAmt, data = LatinoOwnersNotInListing)
    print(summary(lf0))
cat("Investors Sales Price Raw fit ","\n")
    ggplot(LatinoInvestorsNotInListing, aes(x = TaxAmt, y = SalePrice)) +
            geom_smooth(method = "lm") + geom_point() +
            scale_x_continuous(expand = c(0, 0)) + 
            scale_y_continuous(expand = c(0, 0))
    lf0 = lm(SalePrice ~ 0 + TaxAmt, data = LatinoInvestorsNotInListing)
    print(summary(lf0))
```
Notice in above graphs that there seems to be two tax rates, 1.5% and the other at 4%. The high tax seems mostly for properties that sold below $100,000 (for Owners). I will leave why for another time, The high tax rate properties are cut out. A refit of the remaining SalePrice was accomplished. Then all SalePrices were reset to TaxAmt*slope-of-fit.  The intercept has been set to zero.
```{r, echo=FALSE}
ScaleLimit = 30
cat("Owner Sales Price fit with SalePrice/TaxAmt > ",ScaleLimit,"\n")
price = LatinoOwnersNotInListing[with(LatinoOwnersNotInListing,SalePrice/TaxAmt > ScaleLimit),]

    lf0 = lm(SalePrice ~ 0 + TaxAmt, data = price)
    print(summary(lf0))
    ggplot(price, aes(x = TaxAmt, y = SalePrice)) +
            geom_smooth(method = "lm") + geom_point() +
              ylim(00000, 500000) + 
            scale_x_continuous(expand = c(0, 0)) + 
            scale_y_continuous(expand = c(0, 0))
LatinoOwnersNotInListing$SalePrice =
  LatinoOwnersNotInListing$TaxAmt*lf0$coefficients[1]

cat("Owner Sales Price fit with SalePrice/TaxAmt > ",ScaleLimit,"\n")
price = LatinoInvestorsNotInListing[with(LatinoInvestorsNotInListing,
                                         (SalePrice/TaxAmt > ScaleLimit)),] 

    lf0 = lm(SalePrice ~ 0 + TaxAmt, data = price)
    print(summary(lf0))
    ggplot(price, aes(x = TaxAmt, y = SalePrice)) +
            geom_smooth(method = "lm") + geom_point() +
              ylim(00000, 500000) + 
            scale_x_continuous(expand = c(0, 0)) + 
            scale_y_continuous(expand = c(0, 0))
LatinoInvestorsNotInListing$SalePrice =
  LatinoInvestorsNotInListing$TaxAmt*lf0$coefficients[1]


# Sales Price Histogram
yl = "Count"
bw = 10000
cat("Owner  not in ListingsSales Price Histogram using TaxAmt Coeffficent of: ",lf0$coefficients[1],"\n")
# Sale Price
    p = ggplot(LatinoOwnersNotInListing,aes(x=SalePrice)) 
    p +  geom_histogram(binwidth = bw,aes(fill = ..count..)) + 
            xlim(0, 500000) + 
            #ylim(0, 100) +
              stat_bin(binwidth=bw, geom="text",size=3, aes(label=..count..), vjust=-.5) +
            scale_fill_gradient("Count", low = "green", high = "red") +
    labs(title = "Yearly  Latino Owner not in Listings Past Sales Alachua County",x=" Latino Owner Sales Price",y=yl) +
   theme_slanted_axis() 
#
cat("Investor Sales Price Histogram using TaxAmt Coeffficent of: ",lf0$coefficients[1],"\n")
# Sale Price
    p = ggplot(LatinoInvestorsNotInListing,aes(x=SalePrice)) 
    p +  geom_histogram(binwidth = bw,aes(fill = ..count..)) + 
            xlim(0, 500000) + 
            #ylim(0, 100) +
              stat_bin(binwidth=bw, geom="text",size=3, aes(label=..count..), vjust=-.5) +
            scale_fill_gradient("Count", low = "green", high = "red") +
    labs(title = "Yearly  Latino Investor not in Listings Past Sales Alachua County",x=" Latino Investor Sales Price",y=yl) +
   theme_slanted_axis() 
# salesdate / sales price scatter plot
SalesOI = subset(LatinoOwnersNotInListing,
                 LatinoOwnersNotInListing$SaleDate > as.Date("1995-01-01"))

yl = " Latino Owner Sales Price"
bw = 365
SalePriceLimit = 0
tabb = table(SalesOI$SalePrice > SalePriceLimit,SalesOI$PUse)
cat("Sale Price Greater than: ",SalePriceLimit,"\n")
print(round(prop.table(tabb),2))
#
SalePriceLB = subset(SalesOI,(SalesOI$SalePrice > SalePriceLimit))
#png(filename="ListingPastSalesPriceByYear.png", width=8*resol, height=8*resol,res=resol)
p = ggplot(SalePriceLB, aes(x = as.Date(SaleDate), y = SalePrice, color = PUse))
p + geom_point() + 
  geom_abline(intercept = c(50000), slope =0.0, colour = "red", size = 2) +
    geom_abline(intercept = c(75000), slope =0.0, colour = "red", size = 1) +
    geom_abline(intercept = c(100000), slope =0.0, colour = "red", size = 0.5) +
  scale_size_continuous(range = c(3, 8)) +
  scale_y_continuous(limits = c(5000, 300000)) +
  labs(title = "Yearly  Latino Owner not in Listings Past Sales Alachua County",x="Year of Sale",y=yl) +
  theme_slanted_axis()
#dev.off()

bw = 365
SalesOI= subset(LatinoOwnersNotInListing,( 
                  LatinoOwnersNotInListing$SaleDate > as.Date("2000-01-01")))
yl = " Latino Owner Number of Past Sales"
#png(filename="UseListingPastSalesByYear.png", width=8*resol, height=8*resol,res=resol)
p = ggplot(SalesOI, aes(x = as.Date(SaleDate),  fill = PUse))
p + geom_histogram(binwidth = bw) + 
  stat_bin(binwidth=bw, geom="text",size=3, aes(label=..count..), vjust=-1.5) +  
  labs(title = "Yearly Latino Owner not in Listings Past Sales Alachua County",x="Year of Sale",y=yl) +
  theme_slanted_axis()
#dev.off()
bw = 365
yl = "Number of Past Sales"
#png(filename="HomeUseListingPastSalesByYear.png", width=8*resol, height=8*resol,res=resol)
p = ggplot(SalesOI, aes(x = as.Date(SaleDate),  fill = PUse))
p + geom_bar(position="dodge",binwidth = bw) +
#p + geom_histogram(binwidth = bw,aes(fill = ..count..)) + 
#  scale_fill_gradient(yl, low = "clear", high = "white") +
  stat_bin(binwidth=bw, geom="text",size=3, aes(label=..count..), vjust=-1.5) +  
#  scale_size_continuous(range = c(3, 8)) +
  labs(title = "Yearly Latino Owner not in Listings Past Sales Alachua County",x="Year of Sale",y=yl) +
  theme_slanted_axis()
#dev.off()
```

.
