---
title: "CreateRentersMailingList"
author: "Bruce Cottman"
date: "January 21, 2015"
output: html_document
---

This report creates the Gainesville Renter mailing list and various diagnostic graphs.

```{r, echo=FALSE}
### Analyze Last Sale For a given Listing
### 12/15/2014
### Bruce Cottman
###
###
rm(list = ls(all = TRUE)) #CLEAR WORKSPACE
#
invisible(require(XLConnect))
#install.packages("xlsx")
#invisible(require(xlsx))
invisible(require(stringr))
#install.packages("dplyr")
#invisible(require(dplyr))
invisible(require(ggplot2))
#install.packages("lubridate")
invisible(require(lubridate))
#install.packages("bitops")
#invisible(require(bitops))
#invisible(require(RCurl))
#install.packages("scales")
#invisible(require(scales))
#install.packages("ggthemes")
#invisible(library(ggthemes))

#
### graphs theme functions
#theme
theme_slanted_axis = function(){
  theme(plot.title=element_text(size=12,face="bold", vjust=2)) +
    theme(axis.text.x = element_text(angle = 50, vjust=1.0, hjust = 1,face="bold",colour = "black")) +
    theme(axis.text.y = element_text(face="bold",colour = "black")) + 
    theme(axis.text=element_text(size=10),axis.title=element_text(size=10,face="bold"))
}
# DB are usually binary and internal use
readDB <- function(wd,fname){
  setwd(wd)
  cat("Loading database file: ",fname,"\n")
  readRDS(fname)
}
#set target city
TargetCity  = toupper("gainesville")
DateToday ="06072015"
fn = paste0("DB.Alachua.APCA.",DateToday,".rds")
ACPARecords = readDB("~/Documents/Real.Estate/input/DB/",fn)
ACPARecords = ACPARecords[ACPARecords$paddr != "" ,]
ACPARecords = unique(ACPARecords)
#table(ACPARecords$PUse)
# focus on homes, filter out land, vacancies, etc.
pull = c("SINGLE FAMILY",
        "MULTIFAMILY",
        "MISC. RESIDENCE",
        "MOBILE HOME",
        "CONDOMINIUM"
)
ACPARecords = unique(subset(ACPARecords,ACPARecords$PUse %in% pull ))
cat("Loaded ACPA Records:",nrow(ACPARecords),"\n")
# 
# Listings of CITIES OF Alachua county
readListingDB <- function(wdf,fname,wdl){
  setwd(wdf)
  CityList = readWorksheetFromFile(fname)
  ?readWorksheet
  CityList = CityList[complete.cases(CityList),]
  # Read in  listings
  Start =1
  End = nrow(CityList)
  seqi = as.numeric(seq(Start,End))
# ii=seqi[Stat]
  INITFLAG = 0
  for(ii in seqi){
    if(CityList$FILENAME[ii] != "NONE"){
      setwd(wdl)
      prop=readRDS(CityList$FILENAME[ii])
      # INITIALZE
      if(INITFLAG == 0){
        TCityListings = subset(prop,prop$street %in% "0xxxxxyyyzzz")
        INITFLAG = 1
      }
      cat("Number of Listing Records:",nrow(prop)," for City: ",CityList$CITY[ii],"\n")
      # add to total
      TCityListings = rbind(TCityListings,prop)
    } # END IF
  } #END FOR LOOP
  cat("Total Number of Listing Records:",nrow(TCityListings),"\n")
  TCityListings = unique(TCityListings)
  cat("Total Number of UNIQUE Listing Records:",nrow(TCityListings),"\n")
  TCityListings
}

TCityListings = readListingDB("~/Dropbox/Public/Real.Estate/code",
                              "AlachuaCountyCityListingsList.xlsx",
                              "~/Documents/Real.Estate/input/DB/")
cat("Loaded Listings DB:",nrow(TCityListings),"\n")
# no land
TCityListings = unique(subset(TCityListings,TCityListings$type != "LAND"))

# listing address cleanup
TCityListings = TCityListings[TCityListings$street != "100 SLIME ST NO STREET ADDRESS" ,]
cat("Eliminate -100 SLIME ST NO STREET ADDRESS - Records Remaining: ",nrow(TCityListings),"\n")
TCityListings$street = str_replace(TCityListings$street,"TERRACE","TER")
TCityListings$street = str_replace(TCityListings$street," LN"," LANE")
TCityListings$street = str_replace(TCityListings$street," ROAD"," RD")
TCityListings$street = str_replace(TCityListings$street," PLACE"," PL")
TCityListings$street = str_replace(TCityListings$street," BOULEVARD"," BLVD")
TCityListings$street = str_replace(TCityListings$street," # A-"," A")
TCityListings$street = str_replace(TCityListings$street," # B-"," B")

TCityListings$street = str_replace(TCityListings$street," # "," APT ")
TCityListings$street = str_replace(TCityListings$street," AVE AVE"," AVE")
TCityListings$street = str_replace(TCityListings$street," UNIT "," - UNIT ")
TCityListings$street = str_replace(TCityListings$street," SW 45","  SW STATE RD 45")
TCityListings$street = str_replace(TCityListings$street," SW 346","  SW COUNTY RD 346")
TCityListings = subset(TCityListings,str_detect(TCityListings$street,"[0-9]{1,} [NW,NE,SE,SW]"))

cat("Eliminate - NO STREET ADDRESS number - Records Remaining: ",nrow(TCityListings),"\n")
#
# begin Owners,Investers,Renters Loop
# clean up ACPARecords
ACPARecords = ACPARecords[ACPARecords$LocAddr != "",]
cat("Eliminate - null physical ADDRESS number - Records Remaining: ",nrow(ACPARecords),"\n")
# eliminate record columns don't/need in mailing or downstream graphs
ACPARecords$Section = NULL
ACPARecords$Township = NULL
ACPARecords$Range = NULL
ACPARecords$MillageCode = NULL
#ACPARecords$SqFt = NULL
#ACPARecords$HtdSqFt = NULL
ACPARecords$Acres = NULL
ACPARecords$Subdivision = NULL
ACPARecords$AgLandVal = NULL
ACPARecords$MrktLandVal = NULL
ACPARecords$BldgVal = NULL
ACPARecords$MiscVal = NULL
ACPARecords$SOHDeferAmt = NULL
ACPARecords$ExemptAmt = NULL
#ACPARecords$TaxAmt = NULL
ACPARecords$QualUnqual = NULL
ACPARecords$SaleInstrument = NULL
ACPARecords$LocAddr = NULL
ACPARecords$VacImp = NULL
#ACPARecords$Value = NULL
# elimiate any ACPARecords that appear as listing recorda
InListing = intersect(ACPARecords$paddr,TCityListings$street)
NotInListing = ACPARecords[!(ACPARecords$paddr %in% InListing),]
cat("Total Number of Adresses Not in Listing Records:",nrow(NotInListing),"\n")
#create owner and investor record 
# renters, rent from invetors, who do not have tasable as same as physical address
InvestersNotInListing = NotInListing[NotInListing$paddr == NotInListing$maddr,]                  
cat("Investers :physical ADDRESS number does match tax address - Records Remaining: ",nrow(InvestersNotInListing),"\n")
RentersNotInListing = InvestersNotInListing
cat("Renters :physical ADDRESS number does match tax address - Records Remaining: ",nrow(RentersNotInListing),"\n")
OwnersNotInListing = NotInListing[NotInListing$paddr == NotInListing$maddr,]                                  
cat("Owners :physical ADDRESS number does match tax address - Records Remaining: ",nrow(OwnersNotInListing),"\n")
# eliminate record columns don't/need in mailing or downstream graphs
OwnersNotInListing$paddr = NULL
OwnersNotInListing$maddr = NULL
# write out Owner mailing list
setwd("~/Dropbox/realestate.output/sales/")
fln = paste0("OIR.AlachuaCounty.MAILING.",DateToday,".xls")
writeWorksheetToFile(file = fln, data = OwnersNotInListing, sheet = "Owners")
print(getwd())
cat("Saving Owner mailing list to file: ",fln,"\n")
# eliminate record columns don't/need in mailing or downstream graphs
InvestersNotInListing$paddr = NULL
InvestersNotInListing$maddr = NULL
# write out Investers mailing list
writeWorksheetToFile(file = fln, data = InvestersNotInListing, sheet = "Investers")
print(getwd())
cat("Saving Investers mailing list to file: ",fln,"\n")

# eliminate record columns don't/need for renters in mailing or downstream graphs
# this is wrong if other cities than Gainesville for Renters
RentersNotInListing$LastName = "Current Resident"
RentersNotInListing$FirstName1 = NULL
RentersNotInListing$FirstName2 = NULL
RentersNotInListing = within(RentersNotInListing,{Addr1 = LocAddr})
RentersNotInListing$Addr2 = NULL
RentersNotInListing$City = "GAINESVILLE"
RentersNotInListing$State = "FL"
#RentersInACPARecords$Zip = ""
# eliminate record columns don't/need in mailing or downstream graphs
RentersNotInListing$paddr = NULL
RentersNotInListing$maddr = NULL
# write out renter mailing list
writeWorksheetToFile(file = fln, data = RentersNotInListing, sheet = "Renters")
print(getwd())
cat("Saving Renter mailing list to file: ",fln,"\n")
### Graphs
#
cat("Renters not Listing Usage Type:","\n")
tann = table(RentersNotInListing$PUse)
#tann
round(prop.table(tann)*100,1)
#
cat("Renters(actally Investers) live in These Cities:","\n")
tann = sort(table(RentersNotInListing$City),decreasing=TRUE)
bw = 10
leadingCities = head(tann,n=bw)
leadingCities
#
#

yl = "Number in Zip Code"
tann = sort(table(substring(RentersNotInListing$Zip,1,5)),decreasing=TRUE)
bw = 10
leadingZips = head(tann,n=bw)
cat("ACPA Renter zips")

leadingZips
barplot(leadingZips, main="ACPA Invester zips",xlab = "Zip",ylab = yl, 
        col=c("darkblue"), cex.names=0.7)

RentersNotInListing$Zip = NULL
#
bw = 1


#png(filename="Listingbyzip.png", width=11*resol, height=8*resol,res=resol)

#dev.off()
resol = 600
#listings by broker
setwd("~/Documents/Real.Estate/output/graphs/")
yl = "Number of Listings"
bw = 10
tab = sort(table(TCityListings$broker),decreasing=TRUE)
leadingBrokers = head(tab,n=bw)
leadingBrokers
TopBrokerNames = names(leadingBrokers)
maxi = max(tab)
BListings = subset(TCityListings,TCityListings$broker %in% TopBrokerNames )

## set the levels in order we want
BListings = within(BListings, {broker = factor(BListings$broker, 
                              levels=names(sort(table(BListings$broker),decreasing=TRUE)))})

cat("Broker Listing since 2004 \n")
tabb = table(BListings$broker)
round(prop.table(tabb),2)
#png(filename="BrokerListings.png", width=11*resol, height=8*resol,res=resol)
p = ggplot(BListings,aes(x=broker)) 
#p +  geom_histogram(binwidth = bw,aes(fill = ..count..)) +  
  p +  geom_bar(binwidth = bw,aes(fill = ..count..)) +  
  labs(title = "Broker Listings of Gainesville",x="Broker",y=yl) +
#  scale_fill_gradient(yl, low = "red", high = "green") +
  stat_bin(binwidth=bw, geom="text",size=3, aes(label=..count..), vjust=-.5) +
#  scale_y_continuous(limits = c(20, maxi)) +
  theme_slanted_axis() 
#dev.off()
#quarterly Sales since 2010
SalesOI = subset(RentersNotInListing,(RentersNotInListing$SaleDate > as.Date("2000-01-01")))

setwd("~/Documents/Real.Estate/output/graphs/")
yl = "Renter Number of Sales"
bw = 365/4
cat("Quarterly Sales since 2004 \n")
#png(filename="SalesByQuarter.png", width=8*resol, height=8*resol,res=resol)
p = ggplot(SalesOI,aes(x=as.Date(SaleDate))) 
p +  geom_histogram(binwidth = bw,aes(fill = ..count..)) +  
      labs(title = "Quarterly Sales Gainesville",x="Quarter of Sale",y=yl) +
        scale_fill_gradient("Count", low = "red", high = "green") +
        stat_bin(binwidth=bw, geom="text",size=3, aes(label=..count..), vjust=-.5) +
        theme_slanted_axis() 
#dev.off()
# yearly for years gt 2000
cat("Yearly Sales\n")
yl = "Renter Number of Sales"
bw = 365

#png(filename="SalesByYear.png", width=8*resol, height=8*resol,res=resol)
p = ggplot(SalesOI,aes(x=as.Date(SaleDate))) 
p +  geom_histogram(binwidth = bw,aes(fill = ..count..)) +  
        labs(title = "Yearly Sales Gainesville",x="Year of Sale",y=yl) +
        scale_fill_gradient("Count", low = "red", high = "green") +
        stat_bin(binwidth=bw, geom="text",size=3, aes(label=..count..), vjust=-.5) +
        theme_slanted_axis()
#dev.off()
#
cat("Renter Sales Date Catergories by months \n")
tabb = table(RentersNotInListing$SaleDateCategoryLabels)
print(prop.table(tabb))
cat("Percentage of Renter Records with Sales Date > 2002,",round(sum(RentersNotInListing$SaleDate > as.Date("2003-01-01"))
      / nrow(RentersNotInListing),2)*100,"\\%\n")
#
#

SalesOI = subset(RentersNotInListing,RentersNotInListing$SaleDate > as.Date("2000-01-01"))
Adresses = SalesOI
cat("Past Sales of not in Listings\n")
yl = "Number of Past Sales"
bw = 365

#png(filename="ListingPastSalesByYear.png", width=8*resol, height=8*resol,res=resol)
p = ggplot(Adresses,aes(x=as.Date(SaleDate))) 
p +  geom_histogram(binwidth = bw,aes(fill = ..count..)) +  
  labs(title = "Yearly Renter not in Listings Past Sales Gainesville",x="Year of Sale",y=yl) +
  scale_fill_gradient(yl, low = "red", high = "green") +
  stat_bin(binwidth=bw, geom="text",size=3, aes(label=..count..), vjust=-.5) +  
  scale_size_continuous(range = c(3, 8)) +
  theme_slanted_axis()
#dev.off()
cat("Past Sales of  Renters not in Listings by Quarter\n")
yl = "Number of Past Sales"
bw = 365/4

#png(filename="ListingPastSalesByQuarter.png", width=8*resol, height=8*resol,res=resol)
p = ggplot(Adresses,aes(x=as.Date(SaleDate))) 
p +  geom_histogram(binwidth = bw,aes(fill = ..count..)) +  
  labs(title = "Quarterly Renter not in Listings Past Sales Gainesville",x="Year of Sale",y=yl) +
  scale_fill_gradient(yl, low = "red", high = "green") +
  stat_bin(binwidth=bw, geom="text",size=3, aes(label=..count..), vjust=-.5) +  
  scale_size_continuous(range = c(3, 8)) +
  theme_slanted_axis()
#dev.off()
cat("Renters Sales Price Raw fit ","\n")
    ggplot(RentersNotInListing, aes(x = TaxAmt, y = SalePrice)) +
            geom_smooth(method = "lm") + geom_point() +
            scale_x_continuous(expand = c(0, 0)) + 
            scale_y_continuous(expand = c(0, 0))
    lf0 = lm(SalePrice ~ 0 + TaxAmt, data = RentersNotInListing)
    print(summary(lf0))

```
Notice in above graphs that there seems to be two tax rates, 1.2% and the other at 7%. The high tax seems mostly for properties that sold below $100,000 (for Investors). I will leave why for another time, The high tax rate properties are cut out. A refit of the remaining SalePrice was accomplished. Then all SalePrices were reset to TaxAmt*slope-of-fit.  The intercept has been set to zero.
```{r, echo=FALSE}
ScaleLimit = 30
cat("Renter Sales Price fit with SalePrice/TaxAmt > ",ScaleLimit,"\n")
price = RentersNotInListing[with(RentersNotInListing,SalePrice/TaxAmt > ScaleLimit),]

    lf0 = lm(SalePrice ~ 0 + TaxAmt, data = price)
    print(summary(lf0))
    ggplot(price, aes(x = TaxAmt, y = SalePrice)) +
            geom_smooth(method = "lm") + geom_point() +
              ylim(00000, 500000) + 
            scale_x_continuous(expand = c(0, 0)) + 
            scale_y_continuous(expand = c(0, 0))
RentersNotInListing$SalePrice =
  RentersNotInListing$TaxAmt*lf0$coefficients[1]

# Renter Sales Price Histogram
yl = "Count"
bw = 10000
cat("Renter Sales Price Histogram using TaxAmt Coeffficent of: ",lf0$coefficients[1],"\n")
# Sale Price
    p = ggplot(RentersNotInListing,aes(x=SalePrice)) 
    p +  geom_histogram(binwidth = bw,aes(fill = ..count..)) + 
            xlim(0, 500000) + 
            #ylim(0, 100) +
              stat_bin(binwidth=bw, geom="text",size=3, aes(label=..count..), vjust=-.5) +
            scale_fill_gradient("Count", low = "green", high = "red") +
    labs(title = "Yearly Renter Not in Listings Past Sales Gainesville",x=" Latino Owner Sales Price",y=yl) +
   theme_slanted_axis() 
# salesdate / sales price scatter plot
SalesOI = subset(RentersNotInListing,
                 RentersNotInListing$SaleDate > as.Date("1995-01-01"))

yl = " Renter Sales Price"
bw = 365
SalePriceLimit = 0
tabb = table(SalesOI$SalePrice > SalePriceLimit,SalesOI$PUse)
cat("Sale Price Greater than: ",SalePriceLimit,"\n")
print(round(prop.table(tabb),2))
#
SalePriceLB = subset(SalesOI,(SalesOI$SalePrice > SalePriceLimit))
#png(filename="ListingPastSalesPriceByYear.png", width=8*resol, height=8*resol,res=resol)
p = ggplot(SalePriceLB, aes(x = as.Date(SaleDate), y = SalePrice, color = PUse))
p + geom_point() + 
  geom_abline(intercept = c(50000), slope =0.0, colour = "red", size = 2) +
    geom_abline(intercept = c(75000), slope =0.0, colour = "red", size = 1) +
    geom_abline(intercept = c(100000), slope =0.0, colour = "red", size = 0.5) +
  scale_size_continuous(range = c(3, 8)) +
  scale_y_continuous(limits = c(5000, 300000)) +
  labs(title = "Yearly Renter Not in Listings Past Sales Gainesville",x="Year of Sale",y=yl) +
  theme_slanted_axis()
#dev.off()
SalePriceLimit = 50001
cat("Renter count: ",nrow(RentersNotInListing[with(RentersNotInListing, SalePrice < SalePriceLimit),])," below Price: $",SalePriceLimit,"\n")
tabb = table(RentersNotInListing$PUse[with(RentersNotInListing, SalePrice < SalePriceLimit)])
print(tabb)
print(round(prop.table(tabb)*100,1))
SalePriceLimit = 75001
cat("Renter count: ",nrow(RentersNotInListing[with(RentersNotInListing, SalePrice < SalePriceLimit),])," below Price: $",SalePriceLimit,"\n")
tabb=table(RentersNotInListing$PUse[with(RentersNotInListing, SalePrice < SalePriceLimit)])
print(tabb)
print(round(prop.table(tabb)*100,1))
SalePriceLimit = 100001
cat("Renter count: ",nrow(RentersNotInListing[with(RentersNotInListing, SalePrice < SalePriceLimit),])," below Price: $",SalePriceLimit,"\n")
tabb=table(RentersNotInListing$PUse[with(RentersNotInListing, SalePrice < SalePriceLimit)])
print(tabb)
print(round(prop.table(tabb)*100,1))
SalePriceLimit = 125001
cat("Renter count: ",nrow(RentersNotInListing[with(RentersNotInListing, SalePrice < SalePriceLimit),])," below Price: $",SalePriceLimit,"\n")
tabb=table(RentersNotInListing$PUse[with(RentersNotInListing, SalePrice < SalePriceLimit)])
print(tabb)
print(round(prop.table(tabb)*100,1))
bw = 365
SalesOI= subset(RentersNotInListing,( 
                  RentersNotInListing$SaleDate > as.Date("2000-01-01")))
yl = "Renter Number of Past Sales"
#png(filename="UseListingPastSalesByYear.png", width=8*resol, height=8*resol,res=resol)
p = ggplot(SalesOI, aes(x = as.Date(SaleDate),  fill = PUse))
p + geom_histogram(binwidth = bw) + 
  stat_bin(binwidth=bw, geom="text",size=3, aes(label=..count..), vjust=-1.5) +  
  labs(title = "Yearly Renter not in Listings Past Sales Gainesville",x="Year of Sale",y=yl) +
  theme_slanted_axis()
#dev.off()
bw = 365
yl = "Renter Number of Past Sales"
#png(filename="HomeUseListingPastSalesByYear.png", width=8*resol, height=8*resol,res=resol)
p = ggplot(SalesOI, aes(x = as.Date(SaleDate),  fill = PUse))
p + geom_bar(position="dodge",binwidth = bw) +
#p + geom_histogram(binwidth = bw,aes(fill = ..count..)) + 
#  scale_fill_gradient(yl, low = "clear", high = "white") +
  stat_bin(binwidth=bw, geom="text",size=3, aes(label=..count..), vjust=-1.5) +  
#  scale_size_continuous(range = c(3, 8)) +
  labs(title = "Yearly Renter not in  Listings Past Sales Gainesville",x="Year of Sale",y=yl) +
  theme_slanted_axis()
#dev.off()
#
```


