---
title: "PriceFromCMA"
output: html_document
---
A CMA is calculated from historical sales for a given time period, PERIOD, for a given neighborhood ,NBH, for a revalent set of variables.  PERIOD is determined from stability of varibles over that PERIOD.  For Florida, the historical sales data comes for tax apprasor's website - http://www.acpafl.org - Achula county.

Market conditions, MC(j), is obtained by predicting the CMA(j) of listed houses in the neighborhood and comparing to listed prices,   PRICE(j). The PRICE of the target property is 
PRICE(j=T) = CMA(j=T) + MCj=T).

Ideally, PRICE(j)/CMA(j) .vs. time-on-market, TOM, could determine
PRICE(j=T,TOM)
iff TOM data can be found.

```{r,echo = FALSE, message=FALSE}

invisible(require(XLConnect))
#install.packages("xlsx")
invisible(require(xlsx))
invisible(require(stringr))
#install.packages("dplyr")
#require(dplyr)
invisible(require(ggplot2))
#install.packages("lubridate")
require(lubridate)
#install.packages("bitops")
require(bitops)
require(RCurl)
#install.packages("scales")
require(scales)
```


```{r, echo=FALSE}

source("~/Dropbox/Public/Real.Estate/code/Zillow/R/zillow2.R")
source("~/Dropbox/Public/Real.Estate/code/Zillow/R/zillowErrors.R")

#
# READ IN neihborhood files 
# SET WD
setwd("~/Dropbox/Public/Real.Estate/code")
nbh_list = read.xlsx("nbhTargetAddress.xlsx",sheetIndex=1,stringsAsFactors = F)
  #Read in Zillow package
seqi = as.numeric(seq(1,1))
ii=seqi[1]

for(ii in seqi){
  
  cat("Neighborhood: ",nbh_list$Neighborhood[ii],
      " TargetAddress:",nbh_list$Address[ii],
      " TargetZip:", nbh_list$Zip[ii],"\n")
  # SET WD
  setwd("~/Documents/Real.Estate/input/nhhd.taxes")
  nbh = read.csv(nbh_list$File[ii],stringsAsFactors = F)
  nbh$SaleDate = as.Date(nbh$SaleDate, "%m/%d/%y")
  sst = "1615 se 4th st"
  
  zzip = 32641
  dfq = lhcompf(sst, zzip)  
  dfq = lhcompf(nbh_list$Address[ii], nbh_list$Zip[ii])
  dfq$street = toupper(dfq$street)
  dfq$street = str_replace(dfq$street," AVE"," AV")
  #zillow comparables in dfq
  isa = intersect(dfq$street,nbh$LocAddr)
  zc = subset(dfq,dfq$street %in% isa,StringAsFactor=F)
  zc = data.frame(lapply(zc, as.character), stringsAsFactors=FALSE)

  cat("Comparables: ",nrow(zc),"\n")
  for(ij in as.numeric(seq(1,nrow(zc)))){
    cat(zc$street[ij],"\n")
    cat("zestimate value",zc$amount[ij],"\n")
    cat("recent zestimate cost per foot",zc$cpsf[ij],"\n")
    cat("last sold: ")
    print(zc$lastSold[ij])
    cat("last sold Price: ")
    print(zc$lastSoldPrice[ij])
    cat("last sold cost per foot",zc$lcpsf[ij],"\n")
    cat("\n")
    }
  cat("\n")
  
  da = here()-days(c(30,60,90,120,150,180,365))
  nbh180 = subset(nbh,nbh$SaleDate %within% new_interval(da[7],here()))
  nbh180 = subset(nbh180,nbh180$SalePrice > 100)
  if(nrow(nbh180) > 0){
    lf0 = lm(SalePrice ~ 0 + HtdSqFt, data = nbh180)
    
    print(ggplot(nbh180, aes(x = HtdSqFt, y = SalePrice)) +
            geom_point() +
            geom_abline(intercept = 0, slope = lf0$coefficients[[1]][1]) + 
            expand_limits(y=0)
          )
    print(summary(lf0))
    CMA = as.numeric(zc$finishedSqFt[1])*lf0$coefficients[[1]][1]
    cat("Address: ",zc$street[1],zc$zip[1],",Neighborhood Base CMA",CMA,"\n")
    Speed.Of.Sale = c("S","Fast","Medium","Slow","E")
    x = c(.93,.95,1,1.05,1.06)
    CMA.Price = CMA*x
   
    z = data.frame(CMA.Price,x)
    lf0 = lm(CMA.Price ~ 0 + x, data=z)
    summary(lf0)

    print(ggplot(nbh180, aes(x = x, y = CMA.Price)) +
            theme_bw() +
            scale_shape_discrete(solid=T, guide="none") +
            geom_point( shape=c("","F","","S",""),color=c("blue","green","blue","red","blue"),size=20)  +
            geom_abline(intercep=0, slope = lf0$coefficients[[1]][1]) + 
            theme(text = element_text(size=20)) +
#                        ylim(CMA.Price[1]*.95,CMA.Price[3]*1.05) +
#                          scale_y_reverse() +
            scale_y_continuous(labels=dollar) +

#            scale_x_discrete(breaks=c("Fast","Medium","Slow")) +
            labs(x="Speed of Sale", y="Price Based on Past Sales") +
            theme(axis.title.y = element_text(family="serif",face="bold",size=20,angle=90,vjust=0.25)) +
            theme(axis.title.x = element_text(family="serif", face="bold",size=20,angle=00,hjust=0.54, vjust=0)) +

            theme( axis.text.x = element_blank()) 
#            xlim(x[1],x[2]*1.05) 

          )
  } #end if
}  #end for

```

CMA.Price[3]
